Universals of word order result from optimization of grammars for efficient communication

### 2019-04-13 second-draft/3-paper-names-backup.tex
cleaning up

% ,lin2017critical
\dropcap{U}nderstanding and explaining the variation between human languages is the central  goal of Linguistics. % \cite{greenberg1963universals,chomsky1993lectures,givon2009genesis, ...}.
% behaghel1909beziehungen,
% \cite{behaghel1909beziehungen,zipf1949human,greenberg1963universals,lin2017critical} \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}
%An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
Across theoretical paradigms, linguists have hypothesized that language is shaped by efficiency in computation~\cite{chomsky2005three,hauser2002faculty, berwick1984grammatical,hawkins1994performance} and communication~\cite{zipf1949human, Croft:Cruse:2004, Goldberg:2005, pinker1990natural, smith2013linguistic}.
%Role of general computational constraints ~\cite{hauser2002faculty} % berwick2016only
In this work we examine this question from a computational perspective, demonstrating a fully formalized and implemented framework in which a large class of language universals can be explained through the computational optimization of grammars for efficiency. % information-theoretic

@@

Recent work has used an information-theoretic formalization to computationally test these ideas in various domains of language, showing that they predict
both basic statistical properties of languages~\cite{ferreri2003least} and language development~\cite{smith2013linguistic},
and sophisticated aspects of language, such as pragmatic inference \cite{frank2012predicting}, and the distribution of color words~\cite{zaslavsky2018efficient} and kinship categories~\cite{kemp2012kinship} across many languages.
Common to these accounts is the idea of maximizing the amount of information that linguistic forms provide about meanings, while constraining the diversity of forms.

@@

%In Information Theory, the usefulness of a communication channel 
The efficiency of any information-theoretic communication protocol depends on (1) how costly encoding and transmission are, and (2) how precisely messages can be recovered from codes~\cite{shannon1948mathematical}.
%This idea formalizes Zipf's forces of Unification and Diversification
Accordingly, computational tests of functional explanations in linguistics have formalized the overall efficiency of language as a weighted combination of terms representing the amount of information that utterances contain about the underlying messages, and the cost of communication~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge}.

@@

We formalize the cost of a language as its entropy~\cite{ferreri2003least,ferrericancho2007global,futrell2017memory}.
In our setting, this corresponds to the average word-by-word surprisal, the degree to which sentences are unpredictable from the general statistics of the language.
Surprisal has been found to be a highly accurate and general predictor human online processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}, and can be justified  in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
In expectation over all utterances $u$ in a language, the negative surprisal describes the \key{predictability}, or negative entropy, of the utteranes: %this quantity is the entropy: % of the sentences:
%The \key{predictability} of the language is the negative entropy of the sentences:
\begin{equation}
	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
\end{equation}
where the sum runs over all possible sentences $l$ that belong to the language.
In keeping with Zipf's Force of Unification, this quantity describes how homogeneous the language is, i.e., it is larger if the distribution over sentences is concentrated on a smaller number of frequent sentences. %how many distinct sentences there are.

@@

>>>>

% ,lin2017critical
\dropcap{U}nderstanding and explaining the variation between human languages is the central  goal of Linguistics. % \cite{greenberg1963universals,chomsky1993lectures,givon2009genesis, ...}.
% behaghel1909beziehungen,
% \cite{behaghel1909beziehungen,zipf1949human,greenberg1963universals,lin2017critical} \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}
%An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.
Across theoretical paradigms, linguists have hypothesized that language is shaped by efficiency in computation~\cite{chomsky2005three,hauser2002faculty, berwick1984grammatical,hawkins1994performance} and communication~\cite{zipf1949human, Croft:Cruse:2004, Goldberg:2005, pinker1990natural, smith2013linguistic}.
%Role of general computational constraints ~\cite{hauser2002faculty} % berwick2016only
In this work we examine this question from a computational perspective, demonstrating a fully formalized and implemented framework in which a large class of language universals can be explained through the computational optimization of grammars for efficiency. % information-theoretic 
Specifically, we show that efficiency optimization explains  linguistic universals related to word order first documented by Greenberg \cite{greenberg1963universals}.

@@

Recent work has used an information-theoretic formalization to computationally test these ideas in various domains of language, showing that they predict
both basic statistical properties of languages~\cite{ferreri2003least} and language development~\cite{smith2013linguistic},
and sophisticated aspects of language, such as pragmatic inference \cite{frank2012predicting}, and the distribution of color words~\cite{zaslavsky2018efficient} and kinship categories~\cite{kemp2012kinship} across many languages.

@@

%In Information Theory, the usefulness of a communication channel 
The efficiency of any information-theoretic communication protocol depends on (1) how costly encoding and transmission are, and (2) how precisely messages can be recovered from codes~\cite{shannon1948mathematical}.
%This idea formalizes Zipf's forces of Unification and Diversification
Accordingly, computational tests of functional explanations in linguistics have formalized the overall efficiency of language as a weighted combination of terms representing the amount of information that utterances contain about the underlying messages, and the cost of communication~\cite{ferreri2003least,frank2012predicting,zaslavsky2018efficient,kemp2012kinship,regier2015word,goodman2013knowledge}.
Common to these accounts is the idea of maximizing the amount of information that linguistic forms provide about meanings, while constraining the diversity of forms.

@@

We formalize the cost of a language as its entropy~\cite{ferreri2003least,ferrericancho2007global,futrell2017memory}.
In our setting, this corresponds to the average word-by-word surprisal, the degree to which sentences are unpredictable from the general statistics of the language.
Surprisal has been found to be a highly accurate and general predictor of human online processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}, and can be justified  in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
In expectation over all utterances $u$ in a language, the negative surprisal describes the \key{predictability}, or negative entropy, of the utteranes: %this quantity is the entropy: % of the sentences:
%The \key{predictability} of the language is the negative entropy of the sentences:
\begin{equation}
	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
\end{equation}
where the sum runs over all possible sentences $u$ that belong to the language.
In keeping with Zipf's Force of Unification, this quantity describes how homogeneous the language is, i.e., it is larger if the distribution over sentences is concentrated on a smaller number of frequent sentences. %how many distinct sentences there are.

@@

\matmethods{We base our experiments on the Universal Dependencies 2.1 treebanks~\cite{nivre2017universal}. % \footnote{\url{http://universaldependencies.org/}}
We use all languages for which at least one treebank with a training partition was available, a total of 51 languages.
For each language where multiple treebanks with training sets were available, we pooled their training sets; similarly for development sets.
Punctuation was removed.
%We removed punctuation. %, which we operationalized by removing all dependents of \emph{punct} dependencies, and recursively removing their dependents.
